---
title: "Single-cell based deconvolutions of bulk AML transciptomic samples"
author: "E Onur Karakaslar"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Desktop/LUMC/")
```

```{r import libraries and utils, include=TRUE, message=FALSE, warning=FALSE}
library(MuSiC) # deconvolution method
library(xbioc) # required for MusiC to run
library(dplyr) # %>%
library(ggplot2) # nice plots
library(pheatmap) # pheatmaps
library(RColorBrewer) # beatiful colors
library(survminer) # survival analyses (KM-plots)
library(survival) # survival analyses (surv function)

# save figs 
saveFigs = FALSE

# load some utility functions
source("utils/wrangleMat.R") 

# ggplot theme
theme.gg <- theme_bw(base_size = 10) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```


Having established the healthy bone marrow (BM) single-cell map, we can deconvolute
bulk AML patients from different cohorts (TCGA-AML, TARGET, and BEAT-AML).


Importing healthy BM map:
```{r import BM}
# loaded healthy BM map
load("cache/scmap_BM.RDS")

# cell types ordered according to cell counts
cell.count.df <- plyr::count(as.data.frame(C.eset$label))

cell.type.order <- cell.count.df[rev(order(cell.count.df[,2])), 1]
cell.type.order
```


### TCGA
Now we can start our analyses with TCGA-AML data, 
```{r prepare TCGA data, cache=TRUE}
# load count matrix 
count.TCGA <- read.csv("data/Other studies/TCGA/TCGA_ht_seq.csv")

count.TCGA <- wrangleMat(count.TCGA)


# calculate 'stemness score'
meta.TCGA <- read.csv("data/Other studies/TCGA/TCGA_ht_seq_meta.csv", row.names = 1)
meta.TCGA$stemness <- calculateStemness(count.TCGA)


meta.TCGA[meta.TCGA$primary_diagnosis == "Mutated NPM1", "primary_diagnosis"] <- 
    "Acute myeloid leukemia with mutated NPM1"
meta.TCGA[meta.TCGA$primary_diagnosis == "PML-RARa", "primary_diagnosis"] <- 
    "Acute promyelocytic leukaemia, PML-RAR-alpha"
meta.TCGA[meta.TCGA$primary_diagnosis == "inv(16) CBF-beta/MYH11", "primary_diagnosis"] <- 
    "Acute myeloid leukemia, CBF-beta/MYH11"
meta.TCGA[meta.TCGA$primary_diagnosis == "t(8;21) RUNX1-RUNX1T1", "primary_diagnosis"] <- 
    "Acute myeloid leukemia with t(8;21)(q22;q22); RUNX1-RUNX1T1"
meta.TCGA[meta.TCGA$primary_diagnosis == "t(9;11) MLLT3-MLL", "primary_diagnosis"] <- 
    "Acute myeloid leukemia with t(9;11)(p22;q23); MLLT3-MLL"






mut.TCGA <- read.csv("data/Other studies/TCGA/TCGA_mutations_2_oh_meta.csv", row.names = 1)


# primary diagnoses
pdiagnoses.TCGA <- stats::model.matrix(~ 0 + meta.TCGA$primary_diagnosis)
colnames(pdiagnoses.TCGA) <- sub(pattern = "meta.TCGA\\$primary_diagnosis", 
                                 replacement = "", colnames(pdiagnoses.TCGA))


# ens to symbol map
ens2gene <- cinaR::grch38
m <- match(rownames(count.TCGA), ens2gene$ensgene)
mapped.genes <- ens2gene$symbol[m]

# duplicated name/NA/mitochondrial genes
removed.genes <- duplicated(mapped.genes) | is.na(mapped.genes) | grepl("^MT", mapped.genes)

count.TCGA <- count.TCGA[!removed.genes,]
rownames(count.TCGA) <- mapped.genes[!removed.genes]
```

We have prepared the meta / count matrix of TCGA cohort for down-stream analyses.

```{r deconvolute TCGA, cache=TRUE, cache.vars='deconv.TCGA'}
T.eset <- ExpressionSet(assayData = as.matrix(count.TCGA))

# MusiC deconvolution
deconv.TCGA <- music_prop(bulk.eset = T.eset, sc.eset = C.eset, clusters = 'label',
                                            markers = NULL, normalize = FALSE, samples = 'Sample', 
                                            verbose = F)$Est.prop.weighted

# deconvoluted fractions of TCGA samples (first 3 cell types)
head(deconv.TCGA)[,1:3]
```

Now that we have the deconvoluted fractions, we can create a heatmap with patients 
on the rows and cell types on the column, and have a look at it but first, 
we'll assign colors to each deconvoluted cell type for results' reproducibility.
```{r determine colors for each cell type}
cell.type.colors <- c(as.vector(paletteer::paletteer_d("ggthemes::Tableau_20")), "#7d1a1a", "#595959")
names(cell.type.colors) <- (c(colnames(deconv.TCGA), "Mixed Profiles", "Others"))
```


Now, we can have our heatmap:
```{r creating the immune composition heatmap, fig.width=15, fig.height=5}
# TCGA patient annotations 
ann.row.TCGA <- as.data.frame(cbind(pdiagnoses.TCGA, 
                                    Blast.BM = meta.TCGA[,"BM_blasts"],
                                    Blast.PB = meta.TCGA[,"PB_blasts"],
                                    Stemness = meta.TCGA[,"stemness"]))

rownames(ann.row.TCGA) <- colnames(count.TCGA)


# assignment threshold 0.5
# assign each patient to a cluster if they are dominated by a cell type (more than 50%), 
# otherwise they are mixed profiles
clusters.TCGA <- apply(deconv.TCGA > 0.5, 1, function(x){
  ifelse(length(x[x]) > 0, names(x[x]), "Mixed Profiles")
}) 

# assigning colors for reproducibility 
ann.colors <- list(cluster = cell.type.colors[ unique(clusters.TCGA)])

# order patients
new.order.TCGA <- names(clusters.TCGA %>% sort)

breaksList = seq(0,1, by = .001)
pheatmap.colors <- colorRampPalette((brewer.pal(n = 7, name = "Blues")))(length(breaksList))


pmap.TCGA <- pheatmap(t(deconv.TCGA [new.order.TCGA, intersect(cell.type.order, colnames(deconv.TCGA))]), 
         cluster_cols = F, cluster_rows = F, 
         annotation_col = cbind(ann.row.TCGA[new.order.TCGA,], cluster = clusters.TCGA[new.order.TCGA]),
         show_rownames = T, show_colnames = F, 
         color = pheatmap.colors, annotation_colors = ann.colors, 
         breaks = breaksList, border_color = NA)

pmap.TCGA
```


In this heatmap, each row adds up to 1, which composes a single patient. 
Each column represents a different cell type. 9 out of 29 existing cell types were 
removed by `MusiC` since those cell types do not exist in every sample in the BM map.
Therefore, we are left with 20 cell types in total. 

First thing we noticed was, for some patients a particular cell type is over-represented, 
and for others (indicated with dark red, mixed profiles) combinations of few cell types composes the sample.
Therefore, we assigned patients either to their over-represented cell types or to mixed profiles. 

```{r save TCGA deconv hm, include=FALSE}
if (saveFigs){
    pdf("output/Music_TCGA_deconvolution_via_integrated_pseudobulk_simple_clustering.pdf", width = 16, height = 6)
    print(pmap.TCGA)
    dev.off()
}
```

Having the cluster assignments, we can check the survival curves:
```{r survival all clusters, fig.height=4}
temp.TCGA <- cbind(meta.TCGA, cluster = clusters.TCGA)
temp.TCGA$status <- ifelse(temp.TCGA$event == "Alive", 0,1)




# simple clustering
km_trt_fit <- survfit(Surv(overal_survival, status) ~ cluster, data=temp.TCGA)


plot.surv.all.TCGA <- 
    ggsurvplot(km_trt_fit, 
            pval = T,
          
           # Change legends: title & labels
           legend.title = "Cluster",
           legend.labs = gsub("cluster=", "", names(km_trt_fit$strata)), 
          
           palette = cell.type.colors[gsub("cluster=", "", names(km_trt_fit$strata))],
           
           # # Add risk table
           risk.table = TRUE,
           tables.height = 0.25, legend = "none",
           tables.theme = theme_cleantable(),
           
           ggtheme = theme_bw(base_size = 12))

plot.surv.all.TCGA
```


```{r include=FALSE}
if(saveFigs)
{    
    pdf("output/TCGA_Kaplan_Meier_simple_clustering_all.pdf")
    print(plot.surv.all.TCGA, newpage = F)
    dev.off()
}
```

Although there is a difference between cluster assignments, this survival plot looks really messy,
and groups have few samples (e.g. Monocytes having only one patient). Therefore, we redefined the
assignments, and assign the groups that has less than 10 patients to "Others" cluster. Then, replot 
the KM plots once again.

```{r survival redefined clusters}

# redefine groups <15 patients as 'Others'
temp.TCGA <- temp.TCGA %>%  dplyr::add_count(cluster)
temp.TCGA$cluster.refined <- ifelse(temp.TCGA$n > 10, temp.TCGA$cluster, "Others")

km_trt_fit <- survfit(Surv(overal_survival, status) ~ cluster.refined, data= temp.TCGA)


plot.surv.redefined.TCGA <- 
    ggsurvplot(km_trt_fit, 
            pval = T,
          
           # Change legends: title & labels
           legend.title = "Cluster",
           legend.labs = gsub("cluster.refined=", "", names(km_trt_fit$strata)), 
          
           palette = cell.type.colors[gsub("cluster.refined=", "", names(km_trt_fit$strata))],
           
           # # Add risk table
           risk.table = TRUE,
           tables.height = 0.25, legend = "none",
           tables.theme = theme_cleantable(),
           
           ggtheme = theme_bw(base_size = 12))

plot.surv.redefined.TCGA
```

```{r include=FALSE}
if(saveFigs)
{    
    pdf("output/TCGA_Kaplan_Meier_simple_clustering.pdf")
    print(plot.surv.all.TCGA, newpage = F)
    dev.off()
}
```


We can compare primary diagnosis with our assingments as well:
```{r fig.height=4 }
km_trt_fit <- survfit(Surv(overal_survival, status) ~ primary_diagnosis, data=temp.TCGA)

plot.surv.pdiagnosis.TCGA <- 
    ggsurvplot(km_trt_fit, 
           pval = T,
           # surv.median.line = "hv", # Add medians survival
          
           # Change legends: title & labels
          legend.title = "Cluster",
          legend.labs = gsub("primary_diagnosis=", "", names(km_trt_fit$strata)),
           
           palette = "lancet",  
           
           # # Add risk table
           risk.table = TRUE,
           tables.height = 0.25, legend = "none",
           tables.theme = theme_cleantable(),
           
           ggtheme = theme_bw(base_size = 12))

plot.surv.pdiagnosis.TCGA
```

```{r include=FALSE}
if(saveFigs){
    pdf("output/TCGA_Kaplan_Meier_primary_diagnosis.pdf")
    print(plot.surv.pdiagnosis.TCGA, newpage = F)
    dev.off()
}
```

If we group PML-RARa, inv16, and RUNX1-RUNX1T1 
```{r}
temp.TCGA$g1 <- ifelse(temp.TCGA$primary_diagnosis %in% c("Acute promyelocytic leukaemia, PML-RAR-alpha", 
                                                          "Acute myeloid leukemia, CBF-beta/MYH11", 
                                                          "Acute myeloid leukemia with t(8;21)(q22;q22); RUNX1-RUNX1T1"),
                       "Promyelocytes/GMPs (3 groups)", NA)

temp.TCGA[(!(temp.TCGA$g1 %in% "Promyelocytes/GMPs (3 groups)") & 
               temp.TCGA$cluster.refined %in% c("Promyelocytes / GMPs")), "g1"] <-  "Promyelocytes/GMPs (Others)"


km_trt_fit <- survfit(Surv(overal_survival, status) ~ g1, data=temp.TCGA[!is.na(temp.TCGA$g1), ])


plot.promyelocytes.TCGA <- 
    ggsurvplot(km_trt_fit, 
           pval = T,
           # surv.median.line = "hv", # Add medians survival
          
           # Change legends: title & labels
          legend.title = "Cluster",
          legend.labs = gsub("g1=", "", names(km_trt_fit$strata)),
           
           palette = "lancet",  
           
           # # Add risk table
           risk.table = TRUE,
           tables.height = 0.25, legend = "none",
           tables.theme = theme_cleantable(),
           
           ggtheme = theme_bw(base_size = 12))

plot.promyelocytes.TCGA
```
```{r TCGA good prognosis vs others in promyelocytes, fig.width=10}
temp.TCGA$g2 <- ifelse(!is.na(temp.TCGA$g1), temp.TCGA$g1, temp.TCGA$primary_diagnosis)
# covariates <- c("age_at_diagnosis",  "gender", "g1", "stemness", "primary_diagnosis")
# univ_formulas <- sapply(covariates,
#                         function(x) as.formula(paste('Surv(overal_survival, status)~', x)))
#                         
# univ_models <- lapply( univ_formulas, function(x){coxph(x, data = meta.promyelocytes)})
# # Extract data 
# univ_results <- lapply(univ_models,
#                        function(x){ 
#                           x <- summary(x)
#                           p.value<-signif(x$wald["pvalue"], digits=2)
#                           wald.test<-signif(x$wald["test"], digits=2)
#                           beta<-signif(x$coef[1], digits=2);#coeficient beta
#                           HR <-signif(x$coef[2], digits=2);#exp(beta)
#                           HR.confint.lower <- signif(x$conf.int[,"lower .95"], 2)
#                           HR.confint.upper <- signif(x$conf.int[,"upper .95"],2)
#                           HR <- paste0(HR, " (", 
#                                        HR.confint.lower, "-", HR.confint.upper, ")")
#                           res<-c(beta, HR, wald.test, p.value)
#                           names(res)<-c("beta", "HR (95% CI for HR)", "wald.test", 
#                                         "p.value")
#                           return(res)
#                           
#                          })
# res <- t(as.data.frame(univ_results, check.names = FALSE))
# as.data.frame(res)


model <- coxph(Surv(overal_survival, status) ~ age_at_diagnosis + gender + g2 , 
               data = temp.TCGA)


pro.gmp.cox.plot <- ggforest(model, data = temp.TCGA, cpositions = c(0, 0.15, 0.35))
pro.gmp.cox.plot
```



```{r include=FALSE}
if(saveFigs){
    pdf("output/TCGA_Promyelocytes_GMPs_Cox_Regression_Forest.pdf", width = 10)
    pro.gmp.cox.plot
    dev.off()
}
```


```{r NPM1 for different cluster assignments}
temp.TCGA$NPM1.cluster <- ifelse(temp.TCGA$primary_diagnosis == "Acute myeloid leukemia with mutated NPM1",
                                 ifelse(temp.TCGA$cluster == "Promyelocytes / GMPs", 
                                        "NPM1 (Promyelocytes/GMPs)", "NPM1 (Others)"), NA)

km_trt_fit <- survfit(Surv(overal_survival, status) ~ NPM1.cluster, data=temp.TCGA[!is.na(temp.TCGA$NPM1.cluster), ])


plot.promyelocytes.NPM1.TCGA <- 
    ggsurvplot(km_trt_fit, 
           pval = T,
           # surv.median.line = "hv", # Add medians survival
          
           # Change legends: title & labels
          legend.title = "Cluster",
          legend.labs = gsub("NPM1.cluster=", "", names(km_trt_fit$strata)),
           
           palette = "lancet",  
           
           # # Add risk table
           risk.table = TRUE,
           tables.height = 0.25, legend = "none",
           tables.theme = theme_cleantable(),
           
           ggtheme = theme_bw(base_size = 12))

plot.promyelocytes.NPM1.TCGA
```
```{r NOS comparison}
temp.TCGA$NOS.cluster <- ifelse(temp.TCGA$primary_diagnosis == "Acute myeloid leukemia, NOS",
                                 ifelse(temp.TCGA$cluster == "Promyelocytes / GMPs", 
                                        "AML, NOS (Promyelocytes/GMPs)", "AML, NOS (Others)"), NA)


km_trt_fit <- survfit(Surv(overal_survival, status) ~ NOS.cluster, data=temp.TCGA[!is.na(temp.TCGA$NOS.cluster), ])


plot.promyelocytes.NOS.TCGA <- 
    ggsurvplot(km_trt_fit, 
           pval = T,
           # surv.median.line = "hv", # Add medians survival
          
           # Change legends: title & labels
          legend.title = "Cluster",
          legend.labs = gsub("NOS.cluster=", "", names(km_trt_fit$strata)),
           
           palette = "lancet",  
           
           # # Add risk table
           risk.table = TRUE,
           tables.height = 0.25, legend = "none",
           tables.theme = theme_cleantable(),
           
           ggtheme = theme_bw(base_size = 12))

plot.promyelocytes.NOS.TCGA
```



We can also project the [stemness score](https://www.nature.com/articles/nature20598) of the patients per cluster. 
```{r}

p1 <- ggplot(temp.TCGA, aes(cluster, stemness, fill = cluster)) + 
    geom_boxplot() + geom_jitter(width = 0.1) + 
  theme.gg + ylab("Stemness Score") + xlab("") + ggtitle("TCGA") + 
  scale_fill_manual(values = cell.type.colors[unique(temp.TCGA$cluster)])+ 
  labs(fill = "Clusters") + theme(axis.text.x=element_text(angle=45,hjust=1)) 
  
p2 <- ggplot(temp.TCGA, aes(cluster.refined, stemness, fill = cluster.refined)) + 
    geom_boxplot() + geom_jitter(width = 0.1) + 
  theme.gg + ylab("Stemness Score") + xlab("") + ggtitle("TCGA") + 
  scale_fill_manual(values = cell.type.colors[unique(temp.TCGA$cluster.refined)])+ 
  labs(fill = "Refined Clusters") + theme(axis.text.x=element_text(angle=45,hjust=1)) 


print(p1)

print(p2)
```

```{r include=FALSE}
if(saveFigs){
    pdf("output/TCGA_Stemness_vs_Refined_Clusters.pdf", width = 5, height = 5)
    p1 
    p2
    dev.off()
}
```


### LUMC

We have 100 bulk RNA-seq samples from LUMC (in house data). We can deconvolute them 
using the same pipeline. 
```{r deconvoluting LUMC, cache=TRUE}
# count.TCGA <- data.table::fread("data/Studies_Latest/TCGA/TCGA_ht_seq.csv")
# 
# count.TCGA <- as.data.frame(t(count.TCGA))
# 
# colnames(count.TCGA) <- count.TCGA[1,]
# count.TCGA <- count.TCGA[-1,]
# 
# write.csv(count.TCGA, "data/Studies_Latest/TCGA/TCGA_ht_seq.csv")


# count matrix
count.LUMC <- read.csv("data/Other studies/LUMC/LUMC_counts_htseq.csv")

count.LUMC <- wrangleMat(count.LUMC)

meta.LUMC <- read.csv("data/Other studies/LUMC/meta_AML_LUMC.csv", row.names = 1)

meta.LUMC$stemness <- calculateStemness(count.LUMC)
# meta.LUMC.diagnosis <- read.csv("data/Other studies/LUMC/meta_2.csv", sep = ";", row.names = 1)
# meta.LUMC <- cbind(meta.LUMC, meta.LUMC.diagnosis)

# match all primary diagnosis to WHO 2016
meta.LUMC[meta.LUMC$sample_description %in% c("Acute monoblastic/monocytic leukemia",
                                              "Acute monoblastic/monocytic leukemia ",
                                              "Acute myelomonocytic leukemia",
                                              "AML NOS with maturation",
                                              "AML NOS with minimal differentiation",
                                              "AML NOS without maturation"), "primary_diagnosis_refined"] <- 
    "Acute myeloid leukemia, NOS"


meta.LUMC[meta.LUMC$sample_description %in% c("AML with inv(16), CBFB-MYH11"), "primary_diagnosis_refined"] <- 
    "Acute myeloid leukemia, CBF-beta/MYH11"

meta.LUMC[meta.LUMC$sample_description %in% c("AML with mutated NPM1"), "primary_diagnosis_refined"] <- 
    "Acute myeloid leukemia with mutated NPM1"

meta.LUMC[meta.LUMC$sample_description %in% c("AML with inv(3), GATA2, MECOM"), "primary_diagnosis_refined"] <- 
    "Acute myeloid leukemia with inv(3)(q21q26.2) or t(3;3)(q21;q26.2); RPN1-EVI1"


meta.LUMC[meta.LUMC$sample_description %in% c("AML with t(8,21), RUNX1-RUNX1T1"), "primary_diagnosis_refined"] <- 
    "Acute myeloid leukemia with t(8;21)(q22;q22); RUNX1-RUNX1T1"

meta.LUMC[meta.LUMC$sample_description %in% c("AML with t(9,11), KMT2A-MLLT3"), "primary_diagnosis_refined"] <- 
    "Acute myeloid leukemia with t(9;11)(p22;q23); MLLT3-MLL"

meta.LUMC[meta.LUMC$sample_description %in% c("Acute promyelocytic leukemia with t(15,17), PML-RARA"), "primary_diagnosis_refined"] <-
    "Acute promyelocytic leukaemia, PML-RAR-alpha"

meta.LUMC[meta.LUMC$sample_description %in% c("AML with myelodysplasia-related changes"), "primary_diagnosis_refined"] <-
    "Acute myeloid leukemia with myelodysplasia-related changes"


meta.LUMC[meta.LUMC$sample_description %in% c("Therapy-related AML"), "primary_diagnosis_refined"] <-
    "Therapy related myeloid neoplasm"

meta.LUMC[meta.LUMC$sample_description %in% c("AML with t(6,9), DEK-NUP214"), "primary_diagnosis_refined"] <-
    "Acute myeloid leukemia with t(6;9)(p23;q34); DEK-NUP214"



mut.LUMC <- read.csv("data/Other studies/LUMC/LUMC_mutations_oh_meta.csv", row.names = 1)

rownames(mut.LUMC) <- colnames(count.LUMC)

# primary diagnoses
pdiagnoses.LUMC <- stats::model.matrix(~ 0 + meta.LUMC$primary_diagnosis_refined)
colnames(pdiagnoses.LUMC) <- sub(pattern = "meta.LUMC\\$primary_diagnosis_refined", 
                                 replacement = "", colnames(pdiagnoses.LUMC))

# Union FLT3 mutations to get a vector
# mut.LUMC[,"FLT3"] <- ifelse(rowSums(mut.LUMC[,c("FLT3.ITD", "FLT3.TKD")] == 1) > 0, 1, 0)


# ens to symbol map
ens2gene <- cinaR::grch38
m <- match(rownames(count.LUMC), ens2gene$ensgene)
mapped.genes <- ens2gene$symbol[m]

# duplicated name/NA/mitochondrial genes
removed.genes <- duplicated(mapped.genes) | is.na(mapped.genes) | grepl("^MT", mapped.genes)

count.LUMC <- count.LUMC[!removed.genes,]
rownames(count.LUMC) <- mapped.genes[!removed.genes]


T.eset <- ExpressionSet(assayData = as.matrix(count.LUMC))

deconv.LUMC <- music_prop(bulk.eset = T.eset, sc.eset = C.eset, clusters = 'label',
                                            markers = NULL, normalize = FALSE, samples = 'Sample', 
                                            verbose = F)$Est.prop.weighted
```

Now that, we have the deconvoluted fractions we can have the heatmap:
```{r heatmap LUMC, fig.width=15, fig.height=5, cache=TRUE}

ann.row.LUMC <- as.data.frame(cbind(pdiagnoses.LUMC, Blast = meta.LUMC[,"blast_percentage"], 
                                    Stemness = meta.LUMC[,"stemness"] ))

rownames(ann.row.LUMC) <- colnames(count.LUMC)

# assignment threshold
clusters.LUMC <- apply(deconv.LUMC > 0.5, 1, function(x){
  ifelse(length(x[x]) > 0, names(x[x]), "Mixed Profiles")
}) 

new.order.LUMC <- names(clusters.LUMC %>% sort)

# assigning colors for reproducibility 
ann.colors <- list(cluster = cell.type.colors[ unique(clusters.LUMC)])

pmap.LUMC <- pheatmap(t(deconv.LUMC [new.order.LUMC, intersect(cell.type.order, colnames(deconv.LUMC))]), 
         cluster_cols = F, cluster_rows = F, show_rownames = T, show_colnames = F,
         annotation_col = cbind(ann.row.LUMC[new.order.LUMC,], cluster = clusters.LUMC[new.order.LUMC]),
         color = pheatmap.colors, annotation_colors = ann.colors,
         breaks = breaksList, border_color = NA)

pmap.LUMC

```

```{r include=FALSE}
if(saveFigs){
    pdf("output/Music_LUMC_deconvolution_via_integrated_pseudobulk_simple_clustering.pdf", width = 16, height = 5)
    pmap.LUMC
    dev.off()
}
```

```{r stemness vs refined clusters LUMC}
temp.LUMC <- cbind(meta.LUMC, cluster = clusters.LUMC)

# redefine groups <15 patients as 'Others'
temp.LUMC <- temp.LUMC %>%  dplyr::add_count(cluster)
temp.LUMC$cluster.refined <- ifelse(temp.LUMC$n > 10, temp.LUMC$cluster, "Others")

p1 <- ggplot(temp.LUMC, aes(cluster, stemness, fill = cluster)) + 
    geom_boxplot() + geom_jitter(width = 0.1) + 
  theme.gg + ylab("Stemness Score") + xlab("") + ggtitle("LUMC") + 
  scale_fill_manual(values = cell.type.colors[unique(temp.LUMC$cluster)])+ 
  labs(fill = "Clusters") + theme(axis.text.x=element_text(angle=45,hjust=1)) 
  
p2 <- ggplot(temp.LUMC, aes(cluster.refined, stemness, fill = cluster.refined)) + 
    geom_boxplot() + geom_jitter(width = 0.1) + 
  theme.gg + ylab("Stemness Score") + xlab("") + ggtitle("LUMC") + 
  scale_fill_manual(values = cell.type.colors[unique(temp.LUMC$cluster.refined)])+ 
  labs(fill = "Refined Clusters") + theme(axis.text.x=element_text(angle=45,hjust=1)) 


print(p1)
print(p2)
```
```{r include=FALSE}
if(saveFigs){
    pdf("output/LUMC_Stemness_vs_Refined_Clusters.pdf", width = 5, height = 5)
    p1 
    p2
    dev.off()
}
```


### BEAT-AML

```{r BEAT AML deconvolution, cache=TRUE}
count.BEAT <- read.csv("data/Other studies/BEAT/BEAT_AML_ht_seq.csv")

count.BEAT <- wrangleMat(count.BEAT)


meta.BEAT <- read.csv("data/Other studies/BEAT/BEAT_AML_ht_seq_meta.csv", row.names = 1)
meta.BEAT$blasts <- ifelse(grepl("Bone Marrow",meta.BEAT$sample_type), meta.BEAT$BM_blasts, meta.BEAT$PB_blasts)
meta.BEAT$stemness <- calculateStemness(count.BEAT)



meta.BEAT$primary_diagnosis_refined <- ifelse(meta.BEAT$primary_diagnosis %in% c("Acute myeloid leukemia, NOS",
                                                                   "Acute myeloid leukemia with maturation",
                                                                   "Acute myeloid leukemia without maturation",
                                                                   "Acute myeloid leukemia, minimal differentiation",
                                                                   "Acute myelomonocytic leukemia",
                                                                   "Acute monoblastic and monocytic leukemia",
                                                                   "Acute erythroid leukaemia",
                                                                   "Acute megakaryoblastic leukaemia",
                                                                   "Acute basophilic leukemia",
                                                                   "Acute panmyelosis with myelofibrosis"), 
                                              "Acute myeloid leukemia, NOS", meta.BEAT$primary_diagnosis)

meta.BEAT$primary_diagnosis_refined <- ifelse(meta.BEAT$primary_diagnosis_refined %in% 
                                                  c("Atypical chronic myeloid leukemia, BCR/ABL negative",
                                                    "Chronic myelomonocytic leukemia, NOS"), 
                                              "MDS/MPN", meta.BEAT$primary_diagnosis_refined)


meta.BEAT$primary_diagnosis_refined <- ifelse(meta.BEAT$primary_diagnosis_refined %in% 
                                                  c("Essential thrombocythemia"), 
                                              "MPN", meta.BEAT$primary_diagnosis_refined)

meta.BEAT$primary_diagnosis_refined <- ifelse(meta.BEAT$primary_diagnosis_refined %in% 
                                                  c("Mixed phenotype acute leukemia, T/myeloid, NOS"), 
                                              "Acute leukemias of ambiguous lineage", meta.BEAT$primary_diagnosis_refined)


meta.BEAT$primary_diagnosis_refined <- ifelse(meta.BEAT$primary_diagnosis_refined %in% 
                                                  c("Myelodysplastic syndrome, unclassifiable",
                                                    "Refractory anemia with excess blasts",
                                                    "Refractory cytopenia with multilineage dysplasia"), 
                                              "MDS", meta.BEAT$primary_diagnosis_refined)


# Select AML and related neoplasms (WHO 2016 classes)
meta.BEAT[grepl("Acute myeloid leukemia", meta.BEAT$primary_diagnosis_refined) | 
              meta.BEAT$primary_diagnosis_refined %in% c("Acute promyelocytic leukaemia, PML-RAR-alpha",
                                                         "Myeloid sarcoma",
                                                         "Therapy related myeloid neoplasm"),] -> meta.BEAT



mut.BEAT <- read.csv("data/Other studies/BEAT/BEAT_AML_mutations.csv", row.names = 1)

# BEAT has healthy samples, filtering those from the matrices
patients.BEAT <- rownames(meta.BEAT)

count.BEAT <- count.BEAT[,patients.BEAT]
meta.BEAT  <- meta.BEAT [patients.BEAT,]
mut.BEAT   <- mut.BEAT  [patients.BEAT,]


# primary diagnoses
pdiagnoses.BEAT <- stats::model.matrix(~ 0 + meta.BEAT$primary_diagnosis_refined)
colnames(pdiagnoses.BEAT) <- sub(pattern = "meta.BEAT\\$primary_diagnosis_refined",
                                 replacement = "", colnames(pdiagnoses.BEAT))


# ens to symbol map
ens2gene <- cinaR::grch38
m <- match(rownames(count.BEAT), ens2gene$ensgene)
mapped.genes <- ens2gene$symbol[m]

# duplicated name/NA/mitochondrial genes
removed.genes <- duplicated(mapped.genes) | is.na(mapped.genes) | grepl("^MT", mapped.genes)

count.BEAT <- count.BEAT[!removed.genes,meta.BEAT$ID]
rownames(count.BEAT) <- mapped.genes[!removed.genes]

T.eset <- ExpressionSet(assayData = as.matrix(count.BEAT))

deconv.BEAT <- music_prop(bulk.eset = T.eset, sc.eset = C.eset, clusters = 'label',
                                            markers = NULL, normalize = FALSE, samples = 'Sample',
                                            verbose = F)$Est.prop.weighted
```


```{r cache=TRUE, fig.width=16, fig.height=5.5}
ann.row.BEAT <- as.data.frame(cbind(pdiagnoses.BEAT, Blast = meta.BEAT$blasts, Stemness = meta.BEAT$stemness))

rownames(ann.row.BEAT) <- colnames(count.BEAT)


# assignment threshold
clusters.BEAT <- apply(deconv.BEAT > 0.5, 1, function(x){
  ifelse(length(x[x]) > 0, names(x[x]), "Mixed Profiles")
})


# assigning colors for reproducibility
ann.colors <- list(cluster = cell.type.colors[ unique(clusters.BEAT)])

new.order.BEAT <- names(clusters.BEAT %>% sort)

# some of the Blast levels are NA, imputing those using the ones that exists
df.blast <- as.data.frame(cbind(Blast = meta.BEAT$blasts, deconv.BEAT))

fit <- glm(Blast ~ ., data = df.blast, na.action = na.omit )

ann.row.BEAT$Blast.predicted <- predict(fit, newdata = df.blast[,-1])
ann.row.BEAT$Blast.predicted [ann.row.BEAT$Blast.predicted < 0]  <- 0

ann.row.BEAT$Blast.predicted <- ifelse(!is.na(ann.row.BEAT$Blast), ann.row.BEAT$Blast, ann.row.BEAT$Blast.predicted)
ann.row.BEAT <- as.data.frame(subset(ann.row.BEAT, select= -c(Blast)))





pmap.BEAT <- pheatmap(t(deconv.BEAT[new.order.BEAT, intersect(cell.type.order, colnames(deconv.BEAT))]),
             cluster_cols = F, cluster_rows = F, show_rownames = T, show_colnames = F,
             annotation_col = cbind(ann.row.BEAT[new.order.BEAT,], 
                                    cluster = clusters.BEAT[new.order.BEAT]),
             color = pheatmap.colors, 
             annotation_colors = ann.colors,
             breaks = breaksList, border_color = NA,
             gaps_col = cumsum(table(clusters.BEAT[new.order.BEAT])))
pmap.BEAT
```

```{r include=FALSE}
if(saveFigs){
    pdf("output/Music_BEAT_deconvolution_via_integrated_pseudobulk_simple_clustering.pdf", width = 16, height = 5.5)
    pmap.BEAT
    dev.off()
}
```


```{r BEAT immune composition clusters KM plots, fig.height=5}

temp.BEAT <- cbind(meta.BEAT, cluster = clusters.BEAT)

# redefine groups <15 patients as 'Others'
temp.BEAT <- temp.BEAT %>%  dplyr::add_count(cluster)
temp.BEAT$cluster.refined <- ifelse(temp.BEAT$n > 15, temp.BEAT$cluster, "Others")

temp.BEAT$status <- ifelse(temp.BEAT$event == "Alive", 0,ifelse(temp.BEAT$event == "Dead", 1, NA))

km_trt_fit <- survfit(Surv(overal_survival, status) ~ cluster.refined, data=temp.BEAT)


plot.KM.immune.BEAT <- 
    ggsurvplot(km_trt_fit, 
           pval = T,
           # surv.median.line = "hv", # Add medians survival
          
           # Change legends: title & labels
          legend.title = "Cluster",
          legend.labs = gsub("cluster.refined=", "", names(km_trt_fit$strata)),
           
           palette = cell.type.colors[gsub("cluster.refined=", "", names(km_trt_fit$strata))],
           
           # # Add risk table
           risk.table = TRUE,
           tables.height = 0.25, legend = "none",
           tables.theme = theme_cleantable(),
           
           ggtheme = theme_bw(base_size = 12))

plot.KM.immune.BEAT
```

```{r include=FALSE}
if(saveFigs){
    pdf("output/BEAT_Kaplan_Meier_simple_clustering.pdf")
    print(plot.KM.immune.BEAT, newpage = F)
    dev.off()
}
```



```{r}
temp.BEAT$g1 <- ifelse(temp.BEAT$primary_diagnosis %in% c("Acute promyelocytic leukaemia, PML-RAR-alpha", 
                                                          "Acute myeloid leukemia, CBF-beta/MYH11", 
                                                          "Acute myeloid leukemia with t(8;21)(q22;q22); RUNX1-RUNX1T1"),
                       "Promyelocytes/GMPs (3 groups)", NA)



temp.BEAT[(!(temp.BEAT$g1 %in% "Promyelocytes/GMPs (3 groups)") & 
               temp.BEAT$cluster %in% c("Promyelocytes / GMPs")), "g1"] <-  "Promyelocytes/GMPs (Others)"


km_trt_fit <- survfit(Surv(overal_survival, status) ~ g1, data=temp.BEAT[!is.na(temp.BEAT$g1), ])


plot.promyelocytes.BEAT <- 
    ggsurvplot(km_trt_fit, 
           pval = T,
           # surv.median.line = "hv", # Add medians survival
          
           # Change legends: title & labels
          legend.title = "Cluster",
          legend.labs = gsub("g1=", "", names(km_trt_fit$strata)),
           
           palette = "lancet",  
           
           # # Add risk table
           risk.table = TRUE,
           tables.height = 0.25, legend = "none",
           tables.theme = theme_cleantable(),
           
           ggtheme = theme_bw(base_size = 12))

plot.promyelocytes.BEAT
```

```{r BEAT Forest plot, fig.width=14, fig.height=10}
temp.BEAT$g2 <- ifelse(!is.na(temp.BEAT$g1), temp.BEAT$g1, temp.BEAT$primary_diagnosis)


# create the cox regression model for primary diagnosis 
# we exclude 3 groups since they don't have any events
model <- coxph(Surv(overal_survival, status) ~ age_at_diagnosis + gender + g2 , 
               data = temp.BEAT[!temp.BEAT$g2 %in% c("Myeloid sarcoma", 
                                                       "Essential thrombocythemia", 
                                                       "Acute myeloid leukemia without maturation"),])




pro.gmp.cox.plot <- ggforest(model, data = temp.BEAT, cpositions = c(0, 0.12, 0.35))
pro.gmp.cox.plot
```

```{r NPM1 for different cluster assignments BEAT}
temp.BEAT$NPM1.cluster <- ifelse(temp.BEAT$primary_diagnosis_refined == "Acute myeloid leukemia with mutated NPM1",
                                 ifelse(temp.BEAT$cluster == "Promyelocytes / GMPs",
                                        "NPM1 (Promyelocytes/GMPs)", "NPM1 (Others)"), NA)
# 
# temp.BEAT$NPM1.cluster <- ifelse(temp.BEAT$primary_diagnosis == "Acute myeloid leukemia with mutated NPM1",
#                                  temp.BEAT$cluster, NA)

km_trt_fit <- survfit(Surv(overal_survival, status) ~ NPM1.cluster, data=temp.BEAT[!is.na(temp.BEAT$NPM1.cluster), ])


plot.promyelocytes.NPM1.BEAT <- 
    ggsurvplot(km_trt_fit, 
           pval = T,
           # surv.median.line = "hv", # Add medians survival
          
           # Change legends: title & labels
          legend.title = "Cluster",
          legend.labs = gsub("NPM1.cluster=", "", names(km_trt_fit$strata)),
           
           palette = "lancet",  
           
           # # Add risk table
           risk.table = TRUE,
           tables.height = 0.25, legend = "none",
           tables.theme = theme_cleantable(),
           
           ggtheme = theme_bw(base_size = 12))

plot.promyelocytes.NPM1.BEAT
```
```{r BEAT-AML NOS comparison for immune compositions}
## WHO 2016 NOS classification
# AML without maturation 
# AML with maturation 
# Acute myelomonocytic leukemia 
# Acute monoblastic/monocytic leukemia 
# Pure erythroid leukemia 
# Acute megakaryoblastic leukemia 
# Acute basophilic leukemia 
# Acute panmyelosis with myelofibrosis


temp.BEAT$NOS.cluster <- ifelse(temp.BEAT$primary_diagnosis_refined %in% c("Acute myeloid leukemia, NOS"),
                                 ifelse(temp.BEAT$cluster == "Promyelocytes / GMPs", 
                                        "AML, NOS (Promyelocytes/GMPs)", "AML, NOS (Others)"), NA)



km_trt_fit <- survfit(Surv(overal_survival, status) ~ NOS.cluster, data=temp.BEAT[!is.na(temp.BEAT$NOS.cluster), ])


plot.promyelocytes.NOS.BEAT <- 
    ggsurvplot(km_trt_fit, 
           pval = T,
           # surv.median.line = "hv", # Add medians survival
          
           # Change legends: title & labels
          legend.title = "Cluster",
          legend.labs = gsub("NOS.cluster=", "", names(km_trt_fit$strata)),
           
           palette = "lancet",  
           
           # # Add risk table
           risk.table = TRUE,
           tables.height = 0.25, legend = "none",
           tables.theme = theme_cleantable(),
           
           ggtheme = theme_bw(base_size = 12))

plot.promyelocytes.NOS.BEAT
```



### TARGET-AML
```{r TARGET deconvolution, cache=TRUE}
count.TARGET <- read.csv("data/Other studies/TARGET/TARGET_ht_seq.csv")

count.TARGET <- wrangleMat(count.TARGET)

meta.TARGET <- read.csv("data/Other studies/TARGET/TARGET_ht_seq_meta.csv", row.names = 1)
meta.TARGET$stemness <- calculateStemness(count.TARGET)

meta.TARGET$primary_diagnosis_refined <- meta.TARGET$primary_diagnosis

meta.TARGET[meta.TARGET$primary_diagnosis == "Mutated NPM1", "primary_diagnosis_refined"] <- 
    "Acute myeloid leukemia with mutated NPM1"
meta.TARGET[meta.TARGET$primary_diagnosis == "PML-RARa", "primary_diagnosis_refined"] <- 
    "Acute promyelocytic leukaemia, PML-RAR-alpha"
meta.TARGET[meta.TARGET$primary_diagnosis == "inv(16) CBF-beta/MYH11", "primary_diagnosis_refined"] <- 
    "Acute myeloid leukemia, CBF-beta/MYH11"
meta.TARGET[meta.TARGET$primary_diagnosis == "t(8;21) RUNX1-RUNX1T1", "primary_diagnosis_refined"] <- 
    "Acute myeloid leukemia with t(8;21)(q22;q22); RUNX1-RUNX1T1"
meta.TARGET[meta.TARGET$primary_diagnosis == "t(9;11) MLLT3-MLL", "primary_diagnosis_refined"] <- 
    "Acute myeloid leukemia with t(9;11)(p22;q23); MLLT3-MLL"
meta.TARGET[meta.TARGET$primary_diagnosis == "Mutated CEBPA", "primary_diagnosis_refined"] <- 
    "Acute myeloid leukemia with mutated CEBPA"
meta.TARGET[meta.TARGET$primary_diagnosis == "t(6;9) DEK-NUP214", "primary_diagnosis_refined"] <- 
    "Acute myeloid leukemia with t(6;9)(p23;q34); DEK-NUP214"
meta.TARGET[meta.TARGET$primary_diagnosis == "t(3;5) NPM1-MLF1", "primary_diagnosis_refined"] <- 
    "Acute myeloid leukemia with myelodysplasia-related changes"
meta.TARGET[meta.TARGET$primary_diagnosis == "KMT2A-MLLT4", "primary_diagnosis_refined"] <- 
    "Acute myeloid leukemia with t(6;11)(q27;q23.3); MLLT4-KMT2A"
meta.TARGET[meta.TARGET$primary_diagnosis == "KMT2A-MLLT10", "primary_diagnosis_refined"] <- 
    "Acute myeloid leukemia with t(10;11)(p12;q23.3); MLLT10-KMT2A"



mut.TARGET <- read.csv("data/Other studies/TARGET/TARGET_mutations_df_3.csv", row.names = 1)

# primary diagnoses
pdiagnoses.TARGET <- stats::model.matrix(~ 0 + meta.TARGET$primary_diagnosis_refined)
colnames(pdiagnoses.TARGET) <- sub(pattern = "meta.TARGET\\$primary_diagnosis_refined",
                                 replacement = "", colnames(pdiagnoses.TARGET))




# ens to symbol map
ens2gene <- cinaR::grch38
m <- match(rownames(count.TARGET), ens2gene$ensgene)
mapped.genes <- ens2gene$symbol[m]

# duplicated name/NA/mitochondrial genes
removed.genes <- duplicated(mapped.genes) | is.na(mapped.genes) | grepl("^MT", mapped.genes)

count.TARGET <- count.TARGET[!removed.genes,]
rownames(count.TARGET) <- mapped.genes[!removed.genes]


T.eset <- ExpressionSet(assayData = as.matrix(count.TARGET))

deconv.TARGET <- music_prop(bulk.eset = T.eset, sc.eset = C.eset, clusters = 'label',
                                            markers = NULL, normalize = FALSE, samples = 'Sample', 
                                            verbose = F)$Est.prop.weighted
```


```{r fig.width=15, fig.height=5}
ann.row.TARGET <- as.data.frame(cbind(pdiagnoses.TARGET, Blast.BM = meta.TARGET$BM_blasts, 
                                      Blast.PB = meta.TARGET$PB_blasts, 
                                      Stemness = meta.TARGET$stemness))

rownames(ann.row.TARGET) <- colnames(count.TARGET)


# assignment threshold
clusters.TARGET <- apply(deconv.TARGET > 0.5, 1, function(x){
  ifelse(length(x[x]) > 0, names(x[x]), "Mixed Profiles")
}) 



# assigning colors for reproducibility
ann.colors <- list(cluster = cell.type.colors[ unique(clusters.TARGET)])

new.order.TARGET <- names(clusters.TARGET %>% sort)

pmap.TARGET <- pheatmap(t(deconv.TARGET [new.order.TARGET, intersect(cell.type.order, colnames(deconv.TARGET))]),
         cluster_cols = F, cluster_rows = F, show_rownames = T, show_colnames = F,
         annotation_col = cbind(ann.row.TARGET[new.order.TARGET,], cluster = clusters.TARGET[new.order.TARGET]),
         annotation_colors = ann.colors,
         color = pheatmap.colors, breaks = breaksList, border_color = NA, 
         gaps_col = cumsum(table(clusters.TARGET[new.order.TARGET])))

pmap.TARGET
```

```{r include=FALSE}
if(saveFigs){
    pdf("output/TARGET_Kaplan_Meier_simple_clustering.pdf", height = 5, width = 15)
    print(pmap.TARGET, newpage = F)
    dev.off()
}
```


```{r survival analyses TARGET}
temp.TARGET <- cbind(meta.TARGET, cluster = clusters.TARGET)
temp.TARGET$status <- ifelse(temp.TARGET$event == "Dead", 1,0)


temp.TARGET$g1 <- ifelse(temp.TARGET$primary_diagnosis %in% c("inv(16) CBF-beta/MYH11", 
                                                              "t(8;21) RUNX1-RUNX1T1"),
                       "Promyelocytes/GMPs (2 groups)", NA)

temp.TARGET[(!(temp.TARGET$g1 %in% "Promyelocytes/GMPs (2 groups)") & 
               temp.TARGET$cluster %in% c("Promyelocytes / GMPs")), "g1"] <-  "Promyelocytes/GMPs (Others)"


km_trt_fit <- survfit(Surv(overal_survival, status) ~ g1, data=temp.TARGET[!is.na(temp.TARGET$g1), ])


plot.promyelocytes.TARGET <- 
    ggsurvplot(km_trt_fit, 
           pval = T,
           # surv.median.line = "hv", # Add medians survival
          
           # Change legends: title & labels
          legend.title = "Cluster",
          legend.labs = gsub("g1=", "", names(km_trt_fit$strata)),
           
           palette = "lancet",  
           
           # # Add risk table
           risk.table = TRUE,
           tables.height = 0.25, legend = "none",
           tables.theme = theme_cleantable(),
           
           ggtheme = theme_bw(base_size = 12))

plot.promyelocytes.TARGET
```


```{r survival analyses TARGET2, include=FALSE}
# # simple clustering
# temp.TARGET$cluster <- clusters.TARGET
# temp.TARGET <- temp.TARGET %>%  dplyr::add_count(cluster) 
# temp.TARGET$cluster.refined <- ifelse(temp.TARGET$n > 20, temp.TARGET$cluster, "Others")
# 
# km_trt_fit <- survfit(Surv(overal_survival, status) ~ cluster.refined, data=temp.TARGET)
# pdf("output/TARGET_Kaplan_Meier_simple_clustering.pdf")
# p <- ggsurvplot(km_trt_fit, 
#            pval = T,
#            # surv.median.line = "hv", # Add medians survival
#           
#            # Change legends: title & labels
#           legend.title = "Cluster",
#           legend.labs = gsub("cluster.refined=", "", names(km_trt_fit$strata)),
#            
#            palette = "npg",  
#            
#            # # Add risk table
#            risk.table = TRUE,
#            tables.height = 0.25, legend = "none",
#            tables.theme = theme_cleantable(),
#            
#            ggtheme = theme_bw(base_size = 12))
# 
# print(p, newpage = F)
# dev.off()

```

```{r TARGET Promyelocytes forestplot, fig.width=10}
temp.TARGET$g2 <- ifelse(!is.na(temp.TARGET$g1), temp.TARGET$g1, temp.TARGET$primary_diagnosis)

model <- coxph(Surv(overal_survival, status) ~ age_at_diagnosis + gender + g2, 
               data = temp.TARGET)

pro.gmp.cox.plot <- ggforest(model, data = temp.TARGET, cpositions = c(0, 0.15, 0.35))
pro.gmp.cox.plot
```
```{r NPM1 for different cluster assignments TARGET}
temp.TARGET$NPM1.cluster <- ifelse(temp.TARGET$primary_diagnosis == "Mutated NPM1",
                                 ifelse(temp.TARGET$cluster == "Promyelocytes / GMPs",
                                        "NPM1 (Promyelocytes/GMPs)", "NPM1 (Others)"), NA)
# 
# temp.BEAT$NPM1.cluster <- ifelse(temp.BEAT$primary_diagnosis == "Acute myeloid leukemia with mutated NPM1",
#                                  temp.BEAT$cluster, NA)

km_trt_fit <- survfit(Surv(overal_survival, status) ~ NPM1.cluster, data=temp.TARGET[!is.na(temp.TARGET$NPM1.cluster), ])


plot.promyelocytes.NPM1.TARGET <- 
    ggsurvplot(km_trt_fit, 
           pval = T,
           # surv.median.line = "hv", # Add medians survival
          
           # Change legends: title & labels
          legend.title = "Cluster",
          legend.labs = gsub("NPM1.cluster=", "", names(km_trt_fit$strata)),
           
           palette = "lancet",  
           
           # # Add risk table
           risk.table = TRUE,
           tables.height = 0.25, legend = "none",
           tables.theme = theme_cleantable(),
           
           ggtheme = theme_bw(base_size = 12))

plot.promyelocytes.NPM1.TARGET
```


```{r TARGET-AML NOS classification}
temp.TARGET$NOS.cluster <- ifelse(temp.TARGET$primary_diagnosis == "Acute myeloid leukemia, NOS",
                                 ifelse(temp.TARGET$cluster == "Promyelocytes / GMPs", 
                                        "NOS (Promyelocytes/GMPs)", "NOS (Others)"), NA)


km_trt_fit <- survfit(Surv(overal_survival, status) ~ NOS.cluster, data=temp.TARGET[!is.na(temp.TARGET$NOS.cluster), ])


plot.promyelocytes.NOS.TARGET <- 
    ggsurvplot(km_trt_fit, 
           pval = T,
           # surv.median.line = "hv", # Add medians survival
          
           # Change legends: title & labels
          legend.title = "Cluster",
          legend.labs = gsub("NOS.cluster=", "", names(km_trt_fit$strata)),
           
           palette = "lancet",  
           
           # # Add risk table
           risk.table = TRUE,
           tables.height = 0.25, legend = "none",
           tables.theme = theme_cleantable(),
           
           ggtheme = theme_bw(base_size = 12))

plot.promyelocytes.NOS.TARGET
```
